Modeling and analyzing the Corona-virus warning app with the Isabelle Infrastructure framework

### 2020-07-10 paper.tex
Introduction and Related works (incl. graphic)

\institute{Middlesex University London and\\ Technische Universit\"at Berlin\\
\email{f.kammueller@mdx.ac.uk|bialut@gmail.com}
}
\maketitle
\begin{abstract}
We provide a model in the Isabelle Infrastructure framework of the recently published
Corona-virus warning app. The app supports breaking infection chains by informing users
whether they have been in close contact to an infected person. The app has a decentralized
architecture that supports anonymity of users.
We provide a formal model of the existing app with the Isabelle Infrastructure framework
to show up some natural attacks in a very abstract model. We then use the security
refinement process of the Isabelle Infrastructure framework to highlight how the use of
continuously changing ephemeral ids improves the anonymity.
\end{abstract}

\section{Introduction}
\label{sec:intro}
The German Chancellor Angela Merkel has strongly supported the publication of
the mobile phone Corona warning app by publicly proclaiming that the ``Corona
App deserves your trust'' \cite{bundes:20}. Many millions of mobile phone users
in Germany have downloaded the app with 6 million on the first day.
This app is one amongst many similar application that aim at the very important goal
to ``break infection chains'' by providing timely information of users whether they
have been exposed to close contact with a person that has been infected.

The Corona-virus warning app
%has taken a long time to develop being published only on
%16th June 2020. It
was a quite costly project but this was mainly due to the management of
Telekom and SAP being in the driving seat. But the app has been designed with great
attention on privacy:  a distributed architecture \cite{} has been adopted after a long and
heated debate with supporters of a central architecture. The distributed architecture is
based on a very clever distributed application design whereby users' phones are sending
highly anonymized so called ``Ephemeral IDs'' at physical locations via the Bluetooth
protocol. The app saves those IDs of people in close proximity. When at a later date an
infected person reports his infection to a central server, the unique root ID is published
and in the daily check all mobile phones connecting to the central server download
the root IDs of infected people. Since the Ephemeral IDs can be mapped to the root ID
all Ephemeral IDs that have been saved over the last 14 days allow users' phones to
regularly check whether their user has been exposed to an infected person and issue
a warning to the user and recommendation to contact health authorities.
%The warning issued by the Corona warning app entitles to
%having a Corona test done (which at the time of writing is not normally possible).

@@

The Corona-virus app has been produced based on a sophisticated security
concept conceived by experts in the field, %to our knowledge, no formal verification
%
and has strong claims with regard to mathematical support:

``Such a design builds on strong, mathematically provable support for privacy and data protection goals [...]'' \TODOref{DP-3T Whitepaper p. 2} \smallskip\\
%  \NOTE{DP-3T on their design and how wisely chosen it is ;) We have to use this! Something along the l%ines of:}
However, there has been, as of yet, to our knowledge no formal verification been involved.
Even if a ``post-production'' formal specification seems pointless, 
it allows to reveal weak points of the architecture, show that the measures that have
been conceived are suitable to cover those weak points. Thereby, we believe
that our current work is useful to increase the trust in the Corona-virus app necessary for
its wide adoption which in turn is crucial for it to be efficient.

@@

\section{Background and related work}
\label{sec:background}
\subsection{DP-3T and PEPP-PT}
\label{sec:history}
This paper is mainly concerned with the architecture and protocols proposed by the
Decentralized Privacy-Preserving Proximity Tracing project (DP-3T).
The main reason to focus on this particular family of protocols is the
\textit{Exposure Notification Framework}, jointly published by Apple and Google, that adopts
the main core principles of the DP-3T proposal. This API is not only used in the German
Corona-virus warning app but has the potential of being widely adopted in future
app developments that might emerge due to the reach of players like Apple and Google.

@@

Neither DP-3T nor PEPP-PT are synonymous for just one single protocol. Each project endorses
different protocols with unique properties in terms of privacy and data protection.
Yet, on a higher level of abstraction, it seems feasible to distinguish two basic architectures:
protocols as endorsed by PEPP-PT might be characterized as centralized architectures whereas
DP-3T-inspired protocols follow a (more) decentralized approach.\footnote{%
  As we will see, DP-3T involves a central backend server. It is decentralized with regard to
  collection and evaluation of contact information:
  In centralized architectures the server provides a risk scoring services, whereas decentralized
  approaches rely on local risk assessment and, thus, do not need to share contact information with
  the backend.}

\subsubsection{Basic DP-3T protocol: How does it work?}
%  How does it relate to ENF/CWA? (broad strokes)}
%
Upon installation, the app generates secret daily seeds to derive so called \textit{Ephemeral IDs}
(EphIDs) from them. EphIDs are generated locally with cryptographic methods and cannot be connected
to one another but only reconstructed from the secret seed they were derived from.

During normal operation each client broadcasts its EphIDs via Bluetooth whilst scanning for
EphIDs broadcasted by other devices in the vicinity. Collected EphIDs are stored locally along
with associated meta-data such as signal attenuation and date. In DP-3T the contact information
gathered is never shared but only evaluated locally (on the device).

If patients are tested positive on the Corona-virus, they are entitled to upload specific data to a
central backend server. This data is accumulated by the backend server and redistributed to all
clients regularly to provide the means for local risk scoring, i.\,e., determining whether collected
EphIDs match those broadcasted by now-confirmed Corona-virus patients during the last, e.\,g., 14, 
days.

In the most simple (and insecure) protocol proposed by DP-3T
\TODOref{Low-cost decentralized proximity tracing; DP-3T Whitepaper p. 14ff}
this basically translates into publishing the daily seeds used to derive EphIDs from.
Aside from additional security measures,
%\TODOfn[styC]{In short: I want to name it but don't talk about it (since we're not really concerned with% this level of detail). But this sounds just sh...}
like signing content the server provides as a general rule, the protocol implemented by ENF and,
hence, CWA follows this low-cost design.\TODOref{CWA solution architecture ($\rightarrow$ github:CWA); probably ENF specification} DP-3T proposes two other, more sophisticated protocols that improve
privacy and data protection properties to different degrees but are more costly.
%TODOfn[styC]{``complicated'' sounds lame but ``costly'' is, perhaps, to simple/literal} to set up.

>>>>

\institute{Middlesex University London and\\ Technische Universit\"at Berlin\\
\email{f.kammueller@mdx.ac.uk|bialut@gmail.com}
}
\maketitle
\begin{abstract}
We provide a model in the Isabelle Infrastructure framework of the recently published German
Corona-virus warning app. The app supports breaking infection chains by informing users
whether they have been in close contact to an infected person. The app has a decentralized
architecture that supports anonymity of users.
We provide a formal model of the existing app with the Isabelle Infrastructure framework
to show up some natural attacks in a very abstract model. We then use the security
refinement process of the Isabelle Infrastructure framework to highlight how the use of
continuously changing ephemeral ids improves the anonymity.
\end{abstract}

\section{Introduction}
\label{sec:intro}
The German Chancellor Angela Merkel has strongly supported the publication of
the mobile phone Corona warning app (CWA; \cite{cwa:github}) by publicly proclaiming that the ``Corona
App deserves your trust'' \cite{bundes:20}. Many millions of mobile phone users
in Germany have downloaded the app with 6 million on the first day.
CWA is one amongst many similar applications that aim at the very important goal
to ``break infection chains'' by providing timely information to users of whether they
have been in close proximity to a person who tested positive for COVID-19.

%The Corona-virus warning app
%has taken a long time to develop being published only on
%16th June 2020. It
%was a quite costly project but this was mainly due to the management of
%Telekom and SAP being in the driving seat. But the app has been designed with great
The app was designed with great
attention on privacy: a distributed architecture \cite{cwa:arch} has been adopted after a long and
heated debate with supporters of a central architecture. The distributed architecture is
based on a very clever application design whereby users' phones broadcast
highly anonymized so called ``Ephemeral IDs'' at physical locations via the
Bluetooth Low Energy Beacon protocol.
The app saves those IDs of people in close proximity.
When at a later date an infected person reports his infection to a central server,
all clients are provided the means to reconstruct the IDs used by this particular
person's device over the last 14 days.
Therefore, any client can evaluate exposure risks locally and stored contact data has never to be shared.
%
%, the unique root ID is published
%and in the daily check all mobile phones connecting to the central server download
%the root IDs of infected people. Since the Ephemeral IDs can be mapped to the root ID
%all Ephemeral IDs that have been saved over the last 14 days allow users' phones to
%regularly check whether their user has been exposed to an infected person and issue
%a warning to the user and recommendation to contact health authorities.
%The warning issued by the Corona warning app entitles to
%having a Corona test done (which at the time of writing is not normally possible).

@@

The app now in use in Germany has been developed based on a protocol proposed by the DP-3T\cite{dp3t:github} project:
A sophisticated security concept
conceived by experts in the field that has strong claims with regard to mathematical support \cite{dp3t:wp} [p2].
%The Corona-virus app has been produced based on a sophisticated security
%concept conceived by experts in the field %to our knowledge, no formal verification
%%
%and has strong claims with regard to mathematical support:
%
%``Such a design builds on strong, mathematically provable support for privacy and data protection goals [...]'' \TODOref{DP-3T Whitepaper p. 2} \smallskip\\
%  \NOTE{DP-3T on their design and how wisely chosen it is ;) We have to use this! Something along the l%ines of:}
%
However, there has been, as of yet, to our knowledge no formal verification been involved.
Even if a ``post-production'' formal specification seems pointless, 
it allows to reveal weak points of the architecture, show that the measures that have
been conceived are suitable to cover those weak points,
or where and to what extend trade offs have to be made due to inherent vulnerabilities.
Given that the protocol implemented by the framework CWA runs on resembles only the most basic
of the three proposed by the DP-3T project,
a formal verified model that stresses the impact or limits of certain security measures
might help lend some more weight to appeals like \cite{dp3t:impl}
%to Apple and Google
to adopt the more sophisticated protocols.
%
%Thereby, we believe
%that our current work is useful to increase the trust in the Corona-virus app necessary for
%its wide adoption which in turn is crucial for it to be efficient.

@@

\section{Background and related work}
\label{sec:background}
\subsection{DP-3T and PEPP-PT}
\label{sec:history}
We are mainly concerned with the architecture and protocols proposed by the
DP-3T (\textit{Decentralized Privacy-Preserving Proximity Tracing}) project.
The main reason to focus on this particular family of protocols is the
\textit{Exposure Notification Framework} (ENF), jointly published by Apple and Google, that adopts
%the main
core principles of the DP-3T proposal. This API is not only used in the German
Corona-virus warning app but has the potential of being widely adopted in future
app developments that might emerge due to the reach of players like Apple and Google.

@@

Neither DP-3T nor PEPP-PT are synonymous for just one single protocol. Each project endorses
different protocols with unique properties in terms of privacy and data protection.
Yet, on a higher level of abstraction, it seems feasible to distinguish two basic architectures:
protocols as endorsed by PEPP-PT might be characterized as centralized architectures whereas
DP-3T-inspired protocols follow a (more) decentralized approach.\footnote{%
  DP-3T involves a central backend server. It is decentralized with regard to
  the collection and evaluation of contact information:
  In centralized architectures the server provides a risk scoring services, whereas decentralized
  approaches rely on local risk assessment and, thus, do not need to share contact information with
  the backend.}
\\

All DP-3T protocols roughly work the same:
%\subsubsection{Basic DP-3T protocol: How does it work?}
%  How does it relate to ENF/CWA? (broad strokes)}
%
Upon installation, the app generates secret daily seeds to derive \textit{Ephemeral IDs}
(EphIDs) from them. EphIDs are generated locally with cryptographic methods and cannot be connected
to one another but only reconstructed from the secret seed they were derived from.

During normal operation each client broadcasts its EphIDs via Bluetooth whilst scanning for
EphIDs broadcasted by other devices in the vicinity. Collected EphIDs are stored locally along
with associated meta-data such as signal attenuation and date. In DP-3T the contact information
gathered is never shared but only evaluated locally.

If patients test positive for the Corona-virus, they are entitled to upload specific data to a
central backend server. This data is aggregated by the backend server and redistributed to all
clients regularly to provide the means for local risk scoring, i.\,e., determining whether collected
EphIDs match those broadcasted by now-confirmed Corona-virus patients during the last, e.\,g., 14, 
days.

In the most simple (and insecure) protocol proposed by DP-3T
\TODOref{Low-cost decentralized proximity tracing; DP-3T Whitepaper p. 14ff}
this basically translates into publishing the daily seeds used to derive EphIDs from.
The protocol implemented by ENF and,
hence, CWA adopts this low-cost design.\TODOref{CWA solution architecture ($\rightarrow$ github:CWA); probably ENF specification} DP-3T proposes two other, more sophisticated protocols that improve
privacy and data protection properties to different degrees but are more costly to implement.
%Aside from additional security measures,
%\TODOfn[styC]{In short: I want to name it but don't talk about it (since we're not really concerned with% this level of detail). But this sounds just sh...}
%like signing content the server provides as a general rule, the protocol implemented by ENF and,
%hence, CWA follows this low-cost design.\TODOref{CWA solution architecture ($\rightarrow$ github:CWA); probably ENF specification} DP-3T proposes two other, more sophisticated protocols that improve
%privacy and data protection properties to different degrees but are more costly.
%%TODOfn[styC]{``complicated'' sounds lame but ``costly'' is, perhaps, to simple/literal} to set up.

Figure \ref{fig:dp3tprot} illustrates the basic system architecture
along with some of the mitigation measures either implemented in CWA or proposed by DP-3T.

\begin{figure}[htb]
  \begin{center}
    \includegraphics[width=5in]{DP-3T_basic_protocol}
  \end{center}
  \caption{Decentralized privacy-preserving proximity tracing protocol
    (as implemented by CWA)}
  \label{fig:dp3tprot}
\end{figure}
